---
title: "PS270 Final Project"
author: Carley Lane Boudreau
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in possibly exploring data related to consumer behavior. For example, I think exploring the relationship between different social groups on spending or the impacts of different types of advertisement on spending across different social groups could both be interesting.

Does one's age group influence the amount of money they spend on Black Friday?
In this study, I plan to explore consumer behavior specifically on Black Friday and analyze how one's age impacts purchase amount. I hypothesize that younger consumers, on average, will spend more than older shoppers on Black Friday. While the purchase amount reflects actual spending during the event, a shopper’s age group may provide insights into the differences in spending habits and shopping preferences. With this in mind, younger consumers are more likely to be up to date on online deals and more likely to shop online during the event. Additionally, with age, spending patterns may become more conservative or focused on necessity rather than impulse purchases. The explanatory variable is the reported age groups, measured categorically by grouping close age groups together (e.g., a 30-year-old consumer would report 26-35). For example, if the variable were coded =46-50, it would be the equivalent of a consumer reporting to fall somewhere between the ages of 46 and 50. My outcome variable is the purchase amount of each shopper’s transaction(s); the data is numeric, with higher values indicating larger spending. I am somewhat skeptical of the purchase amount values, as some of the numbers seem surprisingly high compared to what I would expect, but the data still follows the logic that higher numbers indicate higher spending. If I observe higher purchase amounts among younger age groups, this would provide support for my hypothesis. However, if I observe lower purchase amounts or no meaningful difference in spending across age groups, this would provide evidence against my hypothesis. 


Adding data from the "Black Friday Sales Data"

```{r}
library(tidyverse)

blfri <-read_csv("train.csv")
blfri

head(blfri)
```


```{r}
blfri_2 <- blfri |>
  group_by(Age) |>
  summarize(avg_purchase = mean(Purchase))
blfri_2

ggplot(blfri_2, aes(x = Age, y = avg_purchase)) +
  geom_point(size = 4, color = "indianred") +
  ylim(8800,9600) +
  labs(
     title = "Average Spending on Black Friday by Age",
    x = "Age Group",
    y = "Average Purchase Amount"
  )
```



